---
title: "R Tools"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# ** Parallel Execution in R**
Using the doParallel library in R, we can set the number of cores of the CPU, which you want your machine to use in while running the model. Keep in mind that assigning all the cores to this process could crash your other processes due to insufficient resources. To be safer, we used the c-2 cores, where c is the number of cores available in your machine.

```{r parallel, eval=FALSE, include=FALSE}
library(doParallel)
### probleme pour restauré en core de base ###

# Find out how many cores are available (if you don't already know)
c =detectCores()
c

# Find out how many cores are currently being used
getDoParWorkers()

# Create cluster with c-2 cores
cl <-makeCluster(c-2)

# Register cluster
registerDoParallel(cl)

# Find out how many cores are being used
getDoParWorkers()

## Example : random forest for binary classifier on fraud on credit card
credit <-read.csv("C:/Users/007/Desktop/Data science with R/R/Dataset/Chapter 9/credit.csv")

# create a random sample for training and test data
set.seed(123)
train_sample <-sample(1000, 900)

# split the data frames
credit_train <-credit[train_sample, ]
credit_test <-credit[-train_sample, ]

##  Building model 
library(randomForest)

#Sequential Execution
system.time(rf_credit_model <-randomForest(credit_train[-17],credit_train$default,ntree =1000))

# In the parallel version of the code, instead of directly using the random forest model with ntree = 1000 parameters, we are going to use the foreach function with %dopar%, so we can split the 1000-decision tree building process into four processes. Each part builds 250 decision trees using the randomForest function.

#Parallel Execution
system.time( rf_credit_model_parallel <-foreach(nt =rep(250,4), 
                                                .combine = combine , 
                                                .packages ='randomForest')
             %dopar%
               randomForest(credit_train[-17],credit_train$default,ntree = nt))

#Shutting down cluster - when you're done, be sure to close #the parallel backend using
stopCluster(cl)
getDoParWorkers()

```

# **Building an R Package** 
=> Beginning Data Science With R p 269

# **Profiling and Optimizing**
=> Beginning Data Science With R p 303


# **Real time data traitement**
    - Manually update model
    - Update in real time
  
  


=> online real-time-based learning methods